---
title: "HW6"
author: "Tianqi Li"
date: "2024-12-02"
output: github_document
---

```{r setup, message = FALSE, warning = FALSE}
library(tidyverse)
library(modelr)
library(mgcv)
library(SemiPar)
set.seed(1)
```

## Problem 1

```{r, message = FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Create the bootstrap samples

```{r}
boot_straps = 
  weather_df |>
  modelr::bootstrap(5000) |>
  mutate(
    strap = map(strap, as_tibble),
    models = map(strap, \(df) lm(tmax~tmin, data = df)),
    glance = map(models, broom::glance), 
    results = map(models, broom::tidy)
  ) |>
  unnest(glance) |>
  select(.id, r.squared, results) |>
  unnest(results) |>
  select(.id, r.squared, term, estimate) |>
  pivot_wider(names_from = term, values_from = estimate) |>
  rename(
    beta0 = `(Intercept)`,
    beta1 = tmin
  ) |>
  mutate(log = log(beta0 * beta1))
```

Plot the distribution

```{r}
boot_straps |>
  ggplot(aes(x = r.squared)) +
  geom_density() +
  ggtitle("Bootstrap Distribution of R-Squared") +
  xlab("R-Squared") +
  ylab("Frequency")

boot_straps |>
  ggplot(aes(x = log)) +
  geom_density() +
  ggtitle("Bootstrap Distribution of log(Beta0 * Beta1)") +
  xlab("log(Beta0 * Beta1)") +
  ylab("Frequency")
```

R-squared distributed between 0.87 to 0.94 with peak at about 0.915.   
log distributed between about 1.90 to 2.10 with peak about 2.02.

```{r}
boot_straps |>
  summarise(
    r2_ci_ll = quantile(r.squared, .025),
    r2_ci_ul = quantile(r.squared, .975),
    log_ci_ll = quantile(log, .025),
    log_ci_ul = quantile(log, .975)
  )
```

## Problem 2

Load the dataset
```{r message = FALSE}
data_url = "https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"

homicide_data = read_csv(data_url)
```

Create the city_state variable and filter the dataset
```{r warning = FALSE}
homicide_df = 
  homicide_data |>
    mutate(
    city_state = paste(city, state, sep = ","),
    solved = ifelse(disposition == "Closed by arrest", 1, 0)
    ) |>
  filter(!(city_state %in% c("Dallas,TX","Phoenix,AZ","Kansas City,MO","Tulsa,AL"))) |>
  filter(victim_race %in% c("White","Black")) |>
  mutate(victim_age = as.numeric(victim_age))
```

Focus on Baltimore city
```{r}
baltimore_test <- 
  homicide_df |>
  filter(city_state == "Baltimore,MD") |>
  glm(solved ~ victim_age + victim_sex + victim_race, family = "binomial", 
      data = _) |>
  broom::tidy(conf.int = TRUE, conf.level = 0.95) |>
  mutate(
    OR = exp(estimate),
    lower = exp(conf.low),
    upper = exp(conf.high)
  ) |>
  filter(term == "victim_sexMale") |>
  select(OR, lower, upper)
baltimore_test |>
  knitr::kable(digits = 3)
```

For each cities
```{r, warning = FALSE}
city_tests <- 
  homicide_df |>
  group_by(city_state) |>
  nest() |>
  mutate(
    fit = map(
      data, ~ glm(solved ~ victim_age + victim_race + victim_sex, data = ., family = binomial())
      ),
    results = map(fit, ~ broom::tidy(.x, conf.int = TRUE, conf.level = 0.95))
  ) |>
  unnest(results) |>
  mutate(
    OR = exp(estimate),
    lower = exp(conf.low),
    upper = exp(conf.high)
  ) |>
  filter(term == "victim_sexMale") |>
  select(city_state, OR, lower, upper)
    
city_tests |>
  knitr::kable(digits = 3)
```

Plot estimated ORs and CIs
```{r, fig.width=8, fig.height=8}
city_tests |>
  ggplot(aes(x = reorder(city_state,desc(OR)), y = OR)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.25, color = "darkgray") +
  labs(
    x = "City",
    y = "Odds Ratio"
  ) +
  coord_flip() +
  theme_minimal()
```

## Problem 3

Load the dataset and clean it
```{r message = FALSE}
bw_df = 
  read_csv("birthweight.csv") |>
  janitor::clean_names() |>
  mutate(
    babysex = factor(babysex, levels = c(1, 2), 
                     labels = c("Male", "Female")),
    malform = factor(malform, levels = c(0, 1), 
                     labels = c("Absent", "Present")),
    frace = factor(frace, levels = c(1, 2, 3, 4, 8, 9), 
                   labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown")),
    mrace = factor(mrace, levels = c(1, 2, 3, 4, 8), 
                   labels = c("White", "Black", "Asian", "Puerto Rican", "Other"))
  )

```

Modeling Process   
Data Cleaning: Variable names were cleaned. Meaningful labels were created for 
factor variables (e.g., babysex, malform) rather than numbers. No missing values 
in the dataset.

```{r}
bw_model = 
  lm(
  bwt ~ babysex + bhead + blength + fincome + frace + gaweeks + malform + momage  + ppbmi + smoken + wtgain, data = bw_df
  )
```

Variables were chosen based on the hypothesized biological and socio-economic 
factors influencing birth weight. My proposed model for birthweight includes the 
following variables: babysex, bhead, blength, fincome, frace, gaweeks, malform, 
momage, ppbmi, smoken, wtgain. Variables like bhead, blength,
and gaweeks directly measure or predict the baby’s growth, making them essential 
for modeling birth weight. Variables like fincome and frace provide insight into 
socio-economic factors that indirectly affect birthweight. Pre-pregnancy BMI, 
and weight gain during pregnancy are key indicators of maternal health and 
directly influence fetal growth.Previous complications increase the risk of 
similar outcomes. These variables are all predictive of birthweight.

Plot residuals vs fitted values
```{r}
bw_df =
  bw_df |>
  modelr::add_predictions(bw_model) |>
  modelr::add_residuals(bw_model)
bw_df |>
  ggplot(aes(x = pred, y = resid)) +
  geom_point() +
  labs(
    x = "Fitted Values",
    y = "Residuals"
  ) +
  theme_minimal()
```

Model comparison
```{r}
cv_df = 
  crossv_mc(bw_df, 100) |>
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  ) |>
  mutate(
    bw_mod = map(train, \(x) lm(bwt ~ babysex + bhead + blength + fincome + frace + gaweeks + malform + momage  + ppbmi + smoken + wtgain, data = x)),
    model1 = map(train, \(x) lm(bwt ~ blength + gaweeks, data = x)),
    model2 = map(train, \(x) lm(bwt ~ bhead * blength * babysex, data = x))
  ) |>
  mutate(
    rmse_bw = map2_dbl(bw_mod, test, rmse),
    rmse_1 = map2_dbl(model1, test, rmse),
    rmse_2 = map2_dbl(model2, test, rmse)
  )

cv_df |>
  select(starts_with("rmse")) |>
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) |>
  ggplot(aes(x = model, y = rmse)) +
  geom_violin()
```

Model 1 uses only blength and gaweeks as predictors and has the highest RMSE, 
indicating it is the least accurate among the three models. Only including two 
variables in the model likely to oversimplify the relationship.

The interaction model performs better than Model 1, as shown by its lower RMSE.
Including interactions between bhead, blength, and babysex adds complexity and 
captures additional variability in birthweight.

The model proposed by myself has the lowest RMSE, suggesting consistent 
performance across cross-validation folds. It performs best likely because 
it incorporates a mix of biological, maternal, and socioeconomic predictors.

